\newpage

\scan{119}

\newpage

% Translation below

\exercise{}
% String' han dig den wijda Mensura, så cavera med något lijtet järn;
% Will han då der på stöta innan til / bruka dig de commetre di spada;
% Det är / cavera åter med en hast/ och gack med korset innan til hoos
% hans swaga / stöt derstädes med secunda til hans högre Axel

If he binds you in wide measure, then disengage with a small amount of iron. If he then seeks to thrust inside, then use the commetre di spada\sidenote{I have been unable to find this term elsewhere, but I suspect it's what's being explained in the following sentence}. That is, disengage again with haste, and go with your cross inside to his weak, thrust thereafter second to his right Shoulder.

\exercise{}
% Lägra dig långt med udden inwertsz / string han då dig utan til / så
% cavera och gif honom jern / der med han får string' dig innan til;
% När han åter på din cavation utan til med ett insteg wil string / så
% giör honom en Mezzia-cavation at du kommer med din Klinga jämpt
% under hans Klinga / då lärer han intet finna din Klinga utan tager
% mistom; Stöt i det samma tempo med en hast Mezzia-cavation / när han
% går din Klinga effter / utan til qvarta öfwer hans högra arm til
% hans Axel a pie fermo.

Position yourself long with the point inwards, if he binds you
outside, then disengage and give him steel, with which he will bind
you inside. When he again on your disengage outside with a step in
seeks to bind, then do him a middle disengage so that your Blade is
straight under his Blade, then he will not find your Blade but instead
miss. Thrust in the same tempo with haste middle disengage, when he
goes after your Blade, utside in fourth over his right arm to his
Shoulder a pie fermo.

% \exercise{Föllier huru man sig retirera skal, och i det samma befrija ifrån Stringering.}
\exercise{Follows how to retreat and in the same free oneself from binds}

% Ligger du i en jämb Tertia / och din Klinga blifwer utan til string'
% / så träd med din wenstra fot något tillbaka / och retirera din
% Klinga tillijka med kroppen / der med du blifwer blott öfwer din
% högra Arm; Så snart din Fiende wil stöta dig i de samma blott /
% Voltera qvartam under hans Klinga till hans wenstre bröst.

If you are in an even Third, and your Blade is bound on the outside, then step with your left foot slightly back, and retreat your Blade and your body, so that you open yourself over the right Arm. As soon as your Enemy seeks to thrust in this opening, Rotate fourthwards under his Blade to his left breast.

\exercise{}
% Retirera dig som tillförende / med din öfwerdel af kroppen / och gif
% dig blott öfwer din högra arm; Wil han då gå utan til / i tanckar at
% så string' dig / så stöt med en hast qvartam förr än han rörer din
% Klinga utan til öfwer hans Arm til wenstre bröstet.

Retreat as follows, with the upper part of your body, and give an
opening over your right arm. If he seeks to go outside, with a thougt
to bind you, then thrust with haste fourth before he touches your
Blade outside over his Arm to the left chest.

\exercise{}
% Der han samma qvarta med Klingan til sin wenstre parerade / så
% passera i en hast under hans Klinga / och stöt med secunda innan til
% hans wenstre bröst / eller stöt secundam öfwer hans wenstre arm.

When he parries your fourth to his left, then pass with haste under
his Blade, and thrust second inside to his left breast, or thrust
scond over his left arm.

\exercise{}
% Hafwer din Adversarius string' utan till / låter du din udd
% derstädes siunka / då är hon åter frij; Wille han din Klinga utan
% till med ett insteg string så gif gran acht uppå tempo / att du i
% det samma caverar och stöter qvart utan till öfwer hans klinga A pie
% fermo till hans högre Axel med ett insteg aff wenstre foten.

If your Adversary has an outside bind, then let your point sink, thus
your blade will be freed. If he seeks to bind your Blade ouside with a
step in then give great attention to tempo, that you in the same
disengage and thrust fourth outside over his blade a pie fermo to his
right Shoulder with a step in by the left foot.

\exercise{}
% Ligger han stilla / när du hafwer försänkt din udd / så gack med din
% klinga åter öfwer sig / at string' honom utan till / men med
% fötterne stat fast / och gif gran acht på han rörelse; Will han i
% det du sänker din udd stöta utan till under din klinga / så stöt med
% liganden secund' contratempo tillijka med honom till hans högra
% sijda / med ett insteg af wenstre foten; Din högre Axel förer du wäl
% fram.

If he lies still, when you have lowered your point ,then go with your
blade again above, to vbind him outside, but with your feet held fast,
and pay great attention to hist movement. If he when you lower your
point thrust under your blade, then thrust with a bind second in
counter-tempo with him to his right side, with a step in of the left
foot. Your right Shoulder bring well forward.

\exercise{}
% Hafwer du åter försänkt din udd / och din Fiende will intet
% förföllia honom / utan förblifwer i sin högd ofwan till / i mening
% at med ett insteg förföllia dig utan till din Klinga / så gif noga
% acht i det han rörer sina fötter / at du då stöter qvart' innan till
% Angulerat med din styrkia i hans swaga till hans högre bröst.

If you have again lowered your point, and your ENemy is not following
it, but remains high and above, meaning to with a step in pursue you
outside your Blade, then pay good attention to when he moves his feet,
that you then thrust fourth inside Angled with your string in his weak
to his right breast.
