\newpage

\scan{084}

\newpage

% Translation below
\exercise{}

% Du kan och när du försänkt din udd i Secunda, och din Fiende wille
% stöta dig utan till öfwer din Klinga in / så parera honom hans stöt
% med styrkan af din Klinga / och medh halfwa hwassan / så tärer du
% förswaga hans Klinga / stöt i dett samma Tert' utan till öfwer hans
% högra Arm: Men der han med sin Klinga tog bak öfwer sig / så passera
% med en hast och stöt Secondam utan til under hans Klinga in. Fig. 5.

You can also when you have lowered your poinmt in Second, and your
Enemy tries to thrust you outside over your Blade in, then parry his
thrust with the string of your Blade, and with the half
sharp\sidenote{Not sure, at all, what this refers to}, you will weaken
his Blade, thrust instantly in Third outside over host right Arm. But
when he whith his Blade takes back over him, then pass with speed and
thrust Second outside in under hist Blade. Fig. 5


\exercise{Against binds}
% \exercise{Emot Stringering.}
% Wille din Advers' string' dig utan til
% så låt din udd med en hast nedersjunka/ wille han förföllia din
% swaga acht uppå i dett samma han förföllier din Klinga / Cavera då
% och stöt innan till Qvart' till hans bröst med ett insteg af högre
% foten: Det han skulle ( dett du stöter Qvart') parera din stöt till
% sin wenstra sinkos / så passera i en hast med din Klinga under hans
% Klinga sinkos och stöt med Secund' till hans inwers Kropp Fig. 16.

If your Adversary desires to bind you on the outside then let your
point quickly sink down, if he pursues your weak, act the instant he
pursues your Blade / Disengage and thrust inwards in Fourth to his
chest with a step in of the right foot: If he were to (in that you
thrusts in Fourth) parry your thrust to his left downwards, then pass
with speed, with your Blade under his Blade and thrust Second to in
inwards Body Fig. 16.



% \exercise{Denna Fig 11 utwijsar en Volte; Och huru man en stringering i ett medel- eller under-läger bemöta skall}
\exercise{This Fig. 11 shows a Volte; And how one a bind in a middle or lower position shall counter}

% Ligger du uti ett högt läger / och han wille string' din Klinga
% innan till / så låt din udd med en hast nedersiunka / Då är din
% Klinga frij; wille han förföllia din swaga / så Cavera i dett samma
% han avancerar och stöt med Tert' utan till öfwer hans Arm till hans
% Bröst med ett utsteg af högre foten; Men der han ( i dett di honom
% öfwer hans halfwa swaga instöter) går med en hast i Qvart' under din
% klinga igenom / och wille stöta Dig en Qvart' innan till, så
% förwexla med en hast utur Tert' i qvart' och voltera qvart' till
% hans inwersz kropp. Fig. 11.

If you are in a high position, and he seeks to bind your Blade inside, then let your tip rapidly drop, then your Blade is free. If he pursues your weak, then Disengage in the instant he advances and thrust Third outside over his Arm to his Chest, with a step out of the right foot. But if he (in the moment over his half weak thrusts) quickly goes to Fourth, through under your blade, and seeks to thrust in Fourth inside, then swich quickly out of Third into fourth and circle aroud fourth to his inside body. Fig 11.
