\newpage

\scan{014}

\newpage
% transcription
% Ligger din Advers' med en jembn utsträckt Arm / och har wändt sin
% Klinga utan till, så gå du med din Klinga ifrå hans wenstra / till
% hans högra / och stringre honom innan til; har han wändt sin udd
% innan til; går du ifrå hans högra til hans wenstra / och stringera
% honom iutan till. Du kan och förr än du stringerar / gå med din udd
% i prospective hans kors / och så med den högre foten fram sättiandes
% iospeso stringera honom / om han då intet moverar sig / du med en
% hast då stoter i samma blott/ När du de under gvardien utan till
% will ligera / så låt din udd uti secunda jämpt under sig siunka /
% och eftter de sccheamater och mutationer de gvardia stringera / som
% altijd elliest med porterter Klinga / hwar med du i en hast som
% tillfälle gifwer / den ene effecten i den andre kan förwända; och
% sätt din foot intet för fort / förr än du har stringerat, dett ware
% då at man hade utag sin Wederparts rörelse / i dett man tillika med
% Klingan och Foten utur den wijda i den [Enga, Euga ??] Mensur avancera kan

% Translation below

If your adversary has an extended arm and has turned his blade
outwads, you should bring your blade from his left to his right and
bind him on the inside. If he's turned his point inwards, you
enter from his right to his left and bind him on the
outside. You can also, before you bind, place your point [i
prospective] his [cross] and with your right foot moving forwards
bind him, if he doesn't move as you thrust into the opening.

When you want to [ligera] on the oustide under the guard, let your
point drop in second and after schemes and changes you can bind
the guard, as always otherwise with [porterer] blade with which you
with haste when an opportunity is given, can turn one effect into the
other; and don't move your foot too fast, until you have bound,
this being that you can with Blade and Foot out of the wide into
narrow Measure advance.

\chap{What Narrow Measure is}

% Stricta, eller then Enga Mensur, är twäggiehanda / såsom Foten
% /Kroppen och Klingan. Mensura Stricta med Foten / är när du har fördt
% foten utur den wijda / och bracht honom i den enga Mensur,
% så at du med blotta öfverböjningen aff din Kropp kan tillräcka din
% Fiende.
% Mensura Stricta med Kroppen / är utan fortsättning din foot /
% allenast med Kroppens öfwersänkning uthur Mens. larga i
% mens. strictam
% Mensura Stricta med Klingan är den distans dijt du med din Klinga
% igenom kropsens öfwerböjning utur den wijda i den ENga Mensur
% bringar.

Narrow, or [Enga] Measure\sidenote{Mensura Stricta}, is two-folded;
Foot, Body and Blade. Narrow Measure with the Foot is when you have
moved the foot out of the wide and brought it into [Enga] Measure, so
a mere leaning of your body is enough to reach your Enemy.

Narrow Measure with the Body is, without moving your foot, simply by
body movement you move out of Long Measure into Strict Measure.

Narrow Measure with the blade is the distance where you bring your
blade from the wide into [Enga] Measure by leaning your body.

% När du nu denna ensur hafwer bekommit (hwilket med stor
% försichtigheet och porterer Klinga motte förättas) kan du på alla
% din Advers Motioner stöta; och der han intet skulle movera sicg /
% ändock igenom den fördel med Contrapsotur aff din [styrkia???] /
% instöta i hans svaga; Men der han skulle uti sin gvardia wara
% temmeligen Copert / år du ändå med din stöt och deportere klinga
% fort. Men om han skull parera / muterar du din effect / och rätar i
% den andre ypningen som kom utaff hans parering.

When you have achieved this measure  (which should be done with great
care and [porterer] blade) you can thrust at all your Adversary's
moves; and should he not move / ths still gives you the advantage  and
through [Contrapostur] from your strong you can thrust into his
weak. But if he should be fairly [Copert] in his guard, you are still
with your thrust and [deportere] blade quickly. But if he should
parry, you can change your move and strike in the opening that his
parry created.

% Retirerar sig din Fiende, när du wil bringa din Klinga i Enga
% mensur, skal du så småningom draga den wenstra foter effter och
% fundera kroppen på honom; Sätt altså den högre foten utan kropsens
% rörelse fram ueder igen så åst du åter i mensura stricta. Skiedde
% din Fiendes retirande i dett du sätter din foor neder är du i
% Larg. motte du straxt fundera din kropp på din högre foot och föllia
% med den wänstra så ofrmert effter, alltid i acht tagandes den Enga
% mensura så procedera. Men merker du att din Fiende ( hwar gång du
% stöter) bryter mensuran skal di intet gå med Furie på honom utan
% låtz förföllia honom och om det gjordes behoff at man där mycket
% frsichtig; En mången ställer sig som han wille fly på det han gifwer
% tilfälleat hans Fiende skal förföllia honom hwar igenom har får
% tempo til att stöta; Är altså intet rådsamt att man förföllier sin
% Fiende med en stark inträngning

If your Enemy retreats when you're bringing your Blade into Narrow
Measure, you should pull your left foot after and center your body on
him. Thus place your right foot without movig your body, and you are
in narrow measure again. If your Enemy's retreat happened as you
lowered your Foot, you are now in Wide. You should quickly shift your
weight to your right foot and let your left follow unnoticeably,
always paying attention that Narrow measure proceeds. But if you
notice that your Enemy (every time that you thrust) breaks measure,
you should not attack him furiously, but pretend to follow him and if
you do, be very careful; Many make it obvious they want to flee, so as
to make his Enemy pursue him, giving him tempo to thrust; It is thus
not advisable to pursue one's Enemy with a forceful entry.