\newpage

\scan{086}

\newpage

% Translation below

% hans sinkos / stöt per Secund utan till under hans Klinga till hans
% högra sijda; En om du skulle gå honom innan till emot så parerade du
% dig sielfwer stöten uppå Kroppen. Fig, 12.

his slowly, theurst Second outside under his Blade to his right
side. But if you were to go at it on the inside you would parry the
thrust yourself on the Body. Fig. 12.

\exercise{}
% Ligger han med kort Arm i secunda eller i den öfwer tert. så
% string. honom innan till men förskrenkter kropp / så snart han går
% under din Klinga till din högra sijda att stöta / så förblif his
% hans klinga / tryck med din klinga hans neder till högra sijda / och
% stöt så med en hast uti ett tempo. utan till med halfwa secunda
% under hans Klinga per passade

If he is positioned with short Arm in the second ör in the upper thrird, then bind him on the inside with lowered body, as soon as he goes to thrust under your Blade to your right side, remain beside his blade, press with your blade his down to the right side, and thrust with haste in a tempo outside with half secondunder his blade per passade.

\exercise{}

% Ligger and i en under Tert så string. honom innan till med
% förhängder qvart. så att din udd kommer innan till hans swaga /
% wille han då innan till under ditt kors med qvart instöta / så gak
% tillijka med honom i qvarta till hans högre höfft; Men giör han en
% volt till ditt öfwerlijf / så giör en Contra-volt utan till öfwer
% hans högra Arm till bröstet; Du kant och sättia din högre fot i dett
% samma tillbaka / och förwända din hand i secunda, gifwandes honom
% stötet i ryggen / och till större säkerheet / för din wenstre hand
% till förskrenkning.

If he is in a lower Third then bind him on the inside qith a hanging fourth so that your point comes inside to his weak, if the then tries to thrust inside under your cross in the fourth, then go similarly in fourth to his right hip. But if he rotates to your upper body, then do a Counter-rotation outside over hist right Arm to the chest. You can also pull your right foot back in that instant, and turn your hand into second, giving him the thrust in the back, and for greater safety, pull your left hand back to close off.

\exercise{}

% Ligera honom utantill med förhängder Secunda, så snart han der med
% qvarta utan till öfwer din klinga wille voltera, så giör en
% Contra-volt under hans höger Arm.

Bind him on the outside with a hanging Second, as soon as he then with
fourth outside over your blade rotates, then do a Counter-rotation
under his right Arm.

\exercise{}

% Guardera dig med lång klinga och string' utan till / eller har han
% string dig så retirera din kropp tillijka med Klingan / att du innan
% till hans Klinga måst i ett underläger kommer / öpna dig derstädes
% innan till / så att din udd något står till din högra sijda; Giör
% han dig en fint innan till / så stöt i dett samma en Qvart' med
% wenstre skenkelen Contratempo yill hans inwerts kropp / Fig. 9. och
% uti stöten gif kroppen wäl fram / wille han aff frij stycken voltera
% öfwer din Arm en qvarta. så stöt under hans klinga secund' till hans
% högra sijda; Men stöter hab qvart' ända fram / så cavera och stöt
% utal til Tert' öfwer hans klinga; String' han innan till / och giör
% dig en stöt ått ditt Ansichte / så passera med seconda under hans
% Klinga.

Guard with a long blade and bind him outside. or if he has bound you
hen retreat your body and Blade, that you get inside his Blade in a
low position, open thenon the inside, so taht your point is slightly
to your right. If ge Feints inside, then trust at the same time a
Fourth with the left ankle Counter-tempo to his inside body,
Fig. 9. and in the thrust give the body well forward, if he with free
piece rotate over your Arm in gourth, then thrust under his blade
second to his right side. But if he thrusts fourth all the way, then
disengage and thrust outide a Third over his blade. If he binds on the
inside and makes a thrust for your Face, then pass with second under
his Blade.
