\newpage

\scan{073}

\newpage

% Translation below

% revolverandes sig till din swaga med sin styrkia / i mening at
% falla bakom din udd / så stöt Tertz. wäl betäckt A. pie fermo under
% eller öfwer hans Klinga / hwarelst han är öppen / och stringera
% honom för effterstöten / retirerar han sig i den samme Paraden, så
% cavera och stringera honom uthan til / eller giör honom der en Fint
% till hans Ansichte; Parerar han öfwer sif / så cavera och stöt under
% hans högre Arm per Secundam. Fig 5.
revolving to your weak with his strong, in order to fall behind your
point, then thrust Third well covered A. pie fermo beneath or above
his Blade, wherever he is open, and bind him for the after-thrust, if
he retreats in this Parry, then disengage and bind him on the outside,
or do him a Feint to his Face; If he parries above him, then disengage
and thrust under his rigth Arm in Second. Fig. 5.


\exercise{}
% Wille han emot denne Finten per Quartam förfalla / syfftandes med
% Secunda till din inwertes Kropp stöta / så låt i en hast din udd
% siunka / och stöt Quart' innan till åht hans högre höft A. pie fermo
% eller med det falska Steget. Här kant du och med blotta Kropsens
% öfwerböjning bruka denne feinten.
If he against this Feint falls to Fourth, with the purpose to thrust
in Second to you inwards Body, let in a hurry your point sink, and
thrust Fourth inside to his right hip A. pie fermo or with the false
Step. Here you can also use this feint with just a bending of the Body.

\exercise{}
% Brukar han sig en tilldragen Parade, och wille med sin styrkia
% upphäfwa din udd eller swaga / så giör en cavade, och stöt med
% halfwa secunda heel uthsträckter innan till hans Lijff / per passada
% eller A. pie fermo. Fig 5.
If he uses a pulled-in Parry, and tries to cancel your point or weak
with his strong, do a disengage, and thrust with half second fully
extended inside to his Midriff, in passada or A. pie fermo. Fig 5.

% \exercise{Chiamater, mer eller utan Retirader giorde.}
% Stringera din Wederpart innan till / och giör dig uthan till öpen
% öfwer din högre Arm / hwilket skeer när du din udd mer än som
% elliest innan till wänder / hwilket och finner sig wid string', der
% udden uthom wanligheeten afwänder utur presensa, eller och med en
% emothålning stöter han med sin Styrka i din swaga / så cavera och
% stöt Quart. innan till.
\exercise{Invitations, with or without Retreats}
If your Opponent binds on the inside, and make yourself open over your
right Arm, which happens when you more than before turn your point
inwards, which is found in the bind, where the point turns more than
common out of presence, or with resistance he thrusts his Strong into
your weak, then disengage and thrust Fourth on the inside.

\exercise{}
% Guardera dig långt / stringera innan till; Eler om du blifwer
% stringerat, så låt din udd i Quarta utan mouvement aff foten /
% allenast rör den freemste leden till din högre sin koos / så att du
% blifwer innan till öfwer Klingan öpen; Har han uthan till
% stringerat, så cavera och giör honom förriga Chiamater; Stöter han i
% det samma Quart' iñan till dit Ansichte / så förfall och stöt Secun'
% under hans Klinga wälbetächt till hans högra Sijda. Fig 21.
Be in a long guard, bind on the inside. Or if you become bound, let
your point in Fourth without movemet of the foot, only move the
frontmost joint to your right bend, so that you become open on the
inside over the Blade. If he has bound on the outside, then disenagge
and do him the previous Invite. If he thrusts instantly in Fourth
inside to your Face, then fall away and thrust Second under his Blade
well covered to his right Side. Fig. 21.
