\newpage

\scan{062}

\newpage

% Translation below

% uti en heel Circel ofwan effter ifrå sin wenstra till sin högra
% muterar / och du den samma effter undfången mensuram i det han går
% med sin Klinga nederåt / stöter du fort till hans högre höfft

a whole Circle above after from his left to his right changes, and you
un the same after received measure in that he goes with his BLade
downwards, thrust quickly to his right hip.

% I synnerhet är denna qvarta till de inwärtes underblotten / emot de
% så wijda innan till Chiamater / så och der til at råka under
% tillijka. De som sänkia eller hafwa sat sig före at råka tillijka /
% hwilker du kan undwijka med det falska steget; Til exempel / när du
% genom den Vantagio med Contrapostur innan eller utan til / och du
% har fått den enga Mensuram / han då innan til giorde en Angulerat
% Chiamata / det du stötte ofwanföre qvarta a: p: fermo fort / han i
% det samma med sin kropp till at råka dig tillijka under din Klinga
% utur präsenza din udd kunde förfalla / skalt du der wijd sänkia din
% udd och voltera qvarta innan till hans underdel / så finner han sig
% bedragen. Men giör ham samma Chiamata in larga / måtte du
% tillförende giöra dig försäkrat för hans udd / och söka så småningom
% at komma hos hans Klinga / eller medelst en Fint få rätta mensuram /
% och sedan hafwa acht på hans rörelse.

In particular is this fourth to the inside lower openings, against the
wide inside invitations, that you then hit below. Those who lowers or
have set themselves to hit thus, which you can avoid with the flase
step. For example, when you with the advantage on the in- or outside,
and you have achieved the measure, he then on the inside made and
Angled invite, that you thrust above in fourth a pie fermo quickly, he
in the same with his body to reach you under your blade, your point
could fall out of presence, you should then lower your point and
rotate fourth on the inside to his lower body, he will find himself
deceived. But if he makes the same Invite large, you must first make
yourself safe from his point, and seek to eventually come to his
Blade, or with a Feint get teh right measure, and then pay attention
to his movements.

% Ähr och så detta stycket till at bruka / när Fienden har aldeles
% satt sig före at taga emot en stöt ofwanföre / allenast han tillijka
% under til får råka / hwarutinnan måste du / om du är utan til /
% allena med kroppsens öfwerböjning utan rörelse af Foten / med
% försänkter udd / genom den Vantagio aff contra postur gå fram för
% dig / och sedan i ett arbete voltera qvartam så lär hans Klinga
% komma innantil din Klinga. Innan til går du med qvarta fram för dig
% till hans högre Axel / i det samma volterar du i ett tempo / udden
% låter du siunka som tilförende under til hans inwärtes kropp / så
% lär du honom i hans förehafwande opprimera / eller giör en Fint der
% emot / och i det at at du upplyfter den högre Foten / see wäl til
% huru man moverar sig / hwar effter du din effect har at mutera: Men
% ligger han stilla / så kommer du honom så när at du din fot geswindt
% sätter neder / i ett tempo stöta och stegen effter lägenhet
% förlängia / och blifwer han (i det han så länge har wäntadt)
% ofelbart råkat / och går du i wijd det samma med en hast åter
% tillbaka igen.

This section is top be used, when the Enemy has decided to recevie a
thrust above, only that he below may reach, where you must, if you are
outside, with only the body's leaning without movement of the Foot,
with lowered point, through the advantage of contraposition go before
yourself, and then in a single move rotate into fourth and his Blade
should be inside your blade. On the inside you go with fourth before
you to his right shoulder, at that instant you rotate in one tempo,
let the point sink below his central body\sidenote{The original
literally has ``internal body''}, you should be able to press down on
what he is doing, or do a Feint against it, and in that you lift your
right Foot, pat good attention to how one moves, after which you can
mutate your effect. But if he is still, you shoudl get so close that
when you swiftly your foot set down, in one tempo thrust and the step
according to cicrumstances lengthen, and he is (in that he so long has
waited) infallibly reached, and you go in that instant with haste back
again.
