\newpage

\scan{101}

\newpage

% Translation below

% rörde foten / woro mycket bättre / allenast med Kropsens
% öfwerböjning; Skulle han i det samma stöta utan till öfwer din
% Klinga / så drag med en hast din Kropp tillbaka igen / och cavera i
% det samma / tag med din Styrka hans Swaga; Förer han sin Hand i
% stöten högt / så stöt med en hast Flanconnade utan till under hans
% högre Arm per qvartam a pie fermo Fig. 16.
moving your foot, that would be better, simply with the leaning of the
body. Were he then to thrust outside, over your Blade, then draw with
haste your Body back, and disengage at the same time, take with your
Strong his Weak. If he brings his Hand high in the thrust, then thrust
with haste {\it Flanconnade}\sidenote{This is a sweeping thrust, where
the point describes a descending arc, with the goal being to end up
under the opponent's arm} outside under his right Arm in fourth a pie
fermo Fig. 16.

\exercise{}
 % Du kant och så / der han skulle fara med sin Arm aff och till i
 % Högden Fig. 18. allenast med din Krops öfwerbörjning / stöta honom
 % i Armen eller Handen.

You can also, were he to change the height of his arm Fig. 18. with
only the leaning of your body, thrust in in the Arm or Hand.

\exercise{}

% Stöter han högt / särdeles till din högre Axel / är intet rådsamt at
% man volterar; En om man då skulle voltera, blefwe man stötter i
% Ryggen: Derföre är det bättre at man står något högt med Kroppen /
% och gifwer sig öfwer Klingan något öpen / så at han hafwer orsak at
% sättia högt an; I det samma tempo han stöter / förwänder du din hand
% i Secunda och förfaller wäl med din Kropp under hans Klinga / och
% passerar med Secunda under hans klinga bort. Fig 22.

If he thrusts high, especially to your right Shoulder, it is not
advised to spin. Since if you spun, you would be thrust in the
Back. Therefore it is better to stand somewhat higher with the Body
and keep somewhat open above the Blade, so that he has reason to aim
high. In the same tempo that he thrusts, turn your hand into Second
and fall well with your Body under his Blade, and pass wirg Second
under his blade away. Fig. 22.

\exercise{}

% Ligger din Advers' i en jämbn tert' fram för dig / så string' honom
% hans Klinga innan till / kastar han då med en hast sin Klinga öfwer
% din / i mening att komma med sin Klinga utå till din KLinga / så
% achta noga när han oplyffter sin Klinga / at du med en hast går
% ifrån din wenstra til din högra under hans Klinga / så lär du åter
% komma innan till hans Klinga; Nät du caverar

If your Adversary lies in a horizontal third in front of you, bind his
Blade on the inside, if he then throws his Blade over yours, seeking
to come with his Blade outside your Blade, then pay good attention
when he lifts his Blade, that you rapidly go from your left to your
right under his Blade, you will again come inside his Blade. When you
disengage
