\newpage

\scan{020}

\newpage

% Translation below

% Contra-Cavation är när du hafwer stringerat din Fiende / och din
% Contrapart will stringera din Klinga igen / att du med dett samma
% går med honom igenom / och kommer på samma ställe / som du
% tillförende låg med din Klinga.

Contra-Cavation is when you have bound your Opponent and your
Counterpart tries to bind your blade again, that you immediately go
through this with him, and end up in the same position, where you
earlier had your blade.

% Ricavation, skeer i dett samma / när du din Advers' will giöra en
% Contra-cavation och du caverar än en gån der på / så finner han sig
% i sing Contra-cavation bedragen.

Ricavantion, happens in that, when your Adversary wants to do a
Contra-cavation and you beat again, so he finds himself in his
Contra-cavation defeated.

% Mezza cavation, är denna/ när du med din udd half igenom går/ och
% med dett samma under hans Klinga blifwer.

Mezza cavation is this, that when you put your point halfway through /
and in that moment go under his Blade.

% Commettere dispada heter / när du caverar. dett är / då du går med
% din udd något högt / och din Fiende will gå till din Klinga / eller
% giöra någon stöt i din blott / du i dett samma går igenom med din
% Klinga / och kommer på samma ställe der du tillförende [waric???]
% hafwer

Commettere dispada is, when you beat, it is when your point is
somewhat high and your Enemy goes for your Blade, or tries to thrust
in an opening, and you in that moment go through with your Blade and
end up in the same place as you previously were.

% Nu på dett din cavation må bättre lyckas eller hafwa fortgång /
% skall du i dett, när du wilt stöta / gå med din udd i en oval fram
% för dig, och med din Klinga hwarken på den ena/ eller den andra
% sijdan/ utan ända fram; En när cavtion är uhte / måste stöten ha
% råkat / elliest lär du intet gärna råka din Fiende a tempo, så kan
% han en håller [ häller? ] när du på detta sätt förhåller dig ( i fall
% han har gifwit tempo ) parera.

So that your beat is more likely to succeed or have progress / you
shall in that when you want to thrust move your point in an oval ahead
of you, and with your Blade neither on one side or the other, but
straight ahead. Whenever beats are out / the thust must have landed /
otherwise you will not reach your Enemy in tempo. Thus he is, when you
proceed in this manner unable to (if he has given tempo)  parry.

% Men med contracavader hade du wäl kunnat befrija dig der ifrå / om
% han har hafft i sinnet att söka stringera din Klinga / blifwer han
% utan all twifwel råkat och stotter/ och är contracavation a tempo
% anbracht / om du ( som tillförende är berört) har welat finna eller
% råka din Fiendes klinga / han åter stökes stringera, går igenom /
% och blifwer med fötterne ferm, och du går tillija med honom igenom
% för än han berört din Klinga.

But with contracavations you would have been able to free yourself, if
he has had in mind to bind your Blade, he will without any doubt be
reached and thrusted / and if contracavation is brought on in tempo /
if you (as has prefiously been discussed) have wanted to find or reach
your Enemy's blade / he again tries to bind, go through / and becoes
with his feet firm, and you will thus with him through before he has
touched your Blade.

% Alltså om han då wille befrija sig ifrå din contracavation, måtte
% han befrija sig med en ricavation i dett samma den andra wille stöta
% med contracavation, har han kunnat råka med en Ricavation.

Thus if he wants to free himself from your contracavation, he must do
so with a ricavation in the same the other wants to thrust with a
contracavtion, he has been able to reach with a Ricavation.

% Comettere dispada kan på fölliande sätt brukat warda / nembligen
% när din Fiende står fast på sine fötter och i en stadig
% stringering / emädan eller i dett samma går igenom / han i en recta
% linea ( för  än du / giort ända på cavation) stöter / och du går än
% på en gång under hans Klinga igenom / så lär du åter komma på samma
% ställe / der du tillförende med din Klinga war; Och d/ar du med din
% styrkia till hans swaga / så har du Salverat dig för hans stöt / och
% lär ha råkat honom i dett samma.

Comettere dispada can in the following wy be used defensively / when
your Enemy stands form on his feet and in a steady bind / in that
moment he goes through / he in a recta linea (before you / has
finished beating) thrusts / and you go through again under his blade /
so you can return to the same place / where you were earlier with your
blade; And then you with your strng to his weak / you have Saved
yourself from his thrust / and should reach him in the same.

% Mezza cavation, blifwer på fölliande sätt förrättat / när din
% Wederpart är dig alt förnär i mens' att du beruchtar dig / dett han
% med sin kropp / för än du har fullbordat din cavation, att hans stöt
% har redan passerat och du

Mezza cavation, is in this fashion executed / when your Opoponent is
too close in measure that you [ beruchtar ??? ] yourswlf / that he
with his body / that when you have completed your cavation, that his
thrust has already passed and you
