\newpage

\scan{052}

\newpage

% Translation below
\chap{How the Four Guards above and below the Blade, without and
within, are thrusted, and with received Measure and Tempo after the
Enemy's posture and motion, used}

\subchaptitle{The thrust in prime and its use}

% Prima blifwer utan till öfver Klingan med hela kroppen stöter / och
% att hon desz swårare är till att parera, måste dett skee med
% försänckter udd. Skickar sig och wäl / att man der (hwarest tertia
% eller secunda wijd högden af Fiendens klinga intet kan stöta eller
% råka) förwänder i prima, emedan derigenom / så wäl utan som innan
% till / med styrkan mehr i högden till hans swaga kommer / kan
% derigenom lätteligen twinga hans Klinga;
Prime is on the outside, above the Blade, with the whole body
thrusted, and to make it much harder to parry, it must be done with a
lowered point. It is well done, that one (whereas third or second, at
the height of the Enemy's blade cannot thrust or reach) turn into
first, through this, on the outside as well as inside, bring the stong
more in the height of his weak, and thus more able to force his Blade.
%                                                                   Skulle han parera med
% secunda öfwer sig / blifwer han då med prima råkat; Och så när du
% hafwer stringerat med tertia utan till / och han derstädes will
% forcera din Klinga / förwänder du i dett samma i prima och låter
% udden wäl siunka.
Should he parry with second above himself, he will then with prime be
reached. And likewise when you have bound with third on the outside,
and he in that place looks to force your blade, you immediately turn
into prime and let the point well sink.
%                               När du hafwer stringerat honom innan till / och
% han i dett samma will cavera, och utan till per secundam stöta
% öfwer din högra Arm / så wändt i dett samma utur Tertia i prima /
% och stott A: p: Fermo förr än han har giordt eller fullbordat
% Cavation; öfwer hans högra Arm  korset wäl uphögt / och kroppen
% jämwäl försänkt / på sådant sätt / blifwe de underblotten
% betäckte.
When you have bound him on the inside, and he in the same wants to
disengage, and thrust on the outside in second above your right Arm,
turn immediately out of Third into first, and thrust a pie Fermo
before he has done or completed the Disengage, over his right arm
cross well lifted, in this way, the lower openings are well covered.
%                  Detta tienar och emot en blott Cavation, så framt du
% honom uti en hög gvard' med Contrapostur innan til något diupt
% kommen åst. Du kan och / när han med en Cavation wil stöta utan till
% öfwer din Klinga tillijka stöta prima med honom.
This also serves agains only a Disengage, as long as you in a high
guard with Contraposture inside somewhat deep reach. You can also,
when he with a Disengage thrusts above your blade thrust him in prime.
%                                                                              Hon låter och
% giöra sig innan til / när du ligger i Tertia eller qvarta, och din
% Fiende stöter dig i din halfwa styrka, då låter du linien fara /
% och förfaller under hans Klinga sin kosz / och stöter prima
% angulerat / med korset wäl högt at winna hans swaga.
This can also be done on the inside, when you are in Third or fourth,
and your Enemy thrusts in your half strong, you let the line go, and
fall under his blade, and thrusts prime angled, with the cross well
high so as to win his weak.

\subchaptitle{Following, the thrust in Second with its use}

% Secunda blifwer stötter / så wäl öfwer / som under Klingan innan
% till / såsom och med kroppen wäl försänkt / der med att du desz
% bättre targer dine blott sinkosz / och hans udd blifwer så mycker
% längre utur presenza och i det
Second is thrusted above as well as below the Blade on the inside,
with the body well lowered, through wich you take your openings lower,
and his point becomes so much more out of presence and in the