\newpage

\scan{118}

\newpage

% Translation below

\exercise{}
% Lägg dig långt i tertia med en jämbd klinga / wille då din Fiende
% med ett insteg innan til string' din klinga / så cavera i det samma
% när han wille taga Mensuram, och förr än han rörer din Klinga / stöt
% innan till hos hans Klinga secundam till hans högre Axel / med ett
% insteg af wenstre foten.

Place yourself in a long third with an even blade, if your Enemy seeks
to bind your blade on the inside with a step in, disengage in the
moment he seeks to take Measure and before he touches your Blade,
thrust second inside by his Blade to his right Shoulder, with a step
in of the left foot.

\exercise{}
% Lägg dig åter som tillförende: Will han dig utan insteg innan til
% string' och har ingen fullkommelig mensuram till stöten / så cavera
% förr än han rörer din Klinga / och giör honom Finten innan till med
% tertia till hans högre Axel; Parerar han til sin wenstra / så wänd i
% qvarta och stöt henne innan till hans bröst med ett insteg aff
% wenstre foten

Place yourself as before. If he binds inside without a step and is not
fully in measure for the thrust, disengage before he touches your
Blade, and do him the inside Feint with third to his right
Shoulder. If he parries to his left, then turn into fourth and thrust
your blade to his chest with a step in of the left foot.

\exercise{}
% Ligger din Wederpart i en hög secunda, och hans udd är något till
% hans högra sijda wänd / så string' honom hans Klinga utan till;
% Caverar han i det samma under din Klinga / cavera tillijka med honom
% / at du kommer åter utan till hans Klinga; Så snart han å nyo går
% under din Klinga igenom / då ryck wäl på honom in / fall med din
% halfwa styrkia på hans halfwa swaga aff Klingan / passera med en
% hast och stöt secund innan till under hans wenstre Arm; Med korset
% bak till wäl högt.
If your opponent is in a high second, and his point is turned slightly
to his right side, then bind his Blade on the outside. If he
disnegages, then disengage with him, so that you again come on the
outside of his Blade. As soon as he again goes through under your
Blade, then jerk close in on him, fall with your half strong on the
half weak of his Blade, pass with haste and thrust second inside under
his left Arm. With the cross back and well high.

\exercise{}
% Ligger en uti jämbd tertia för dig / så stringera honom hans Klinga
% innan till / går han i det under din Klinga igenom / så gif gran
% acht på tempo, wijd samma han caverar under din Klinga / at du med
% din styrkia då batterar i hans swaga / stöt med en hast qvartam utan
% till öfwer hans wenstra Arm till bröstet med ett insteg af högre
% foten.
If he is in an even third before you, then bind his Blade inside, if
he in that goes through under your Blade, pay good attention to tempo,
and in the same as he disengages under your Blade, that you with your
strong beat in his weak, thrust with haste fourth outside over his
left Arm to the chest with an step in of the right foot.

\exercise{}
% Stringera honom innan til; Caverar han / och wille dig igen innan til
% med ett insteg string / så Contra-Cavera och stöt i det samma med
% Contra-Cavation innan till secundam til hans högre Axel med ett
% insteg af wenstre foten.
Bind him inside. If he disengages, and seek to bind you again with a
step in, then Counter-Disengage and thrust immediately with the
Counter-Disengage inside second to his right Shoulder with a step in
of the left foot.

\exercise{}
% På ett annat sätt: String' honom utan til / går han i dett samma
% igenom / och wil dig förutan inträdning tilbakas[???] utan til
% stringera Contra cavera och string honom utan til; Caverar han åter
% igen / så stöt wijd begynnelsen i cavat innan til med secunda; Udden
% blifwer försänkt innan til hans bröst.
In another way: Bind him outside, if he immediately go through, and
wants without a step in back outside bind, Counter-disengage and bind
him outside. Iv he disnegages again, then thrust at the beginning of
the disengagement inside with second. The point remains lowered,
inside to his chest.

\exercise{}
% Lägg dig långt med udden något inwersz / string han dig utan til /
% så cavera; contra caverar han och träder med en string' i mensuram /
% Ricavera med en hast / och stöt med Ricavation utan til qvartam
% öfwer hans Klinga; träder han med contra-cavation intet för sig / så
% giör honom med Ricavation utan til i qvarta en Fint med järn / och
% gif gran acht uppå huru han parerar.
Lie in a long position with the point slightly inside, if he binds you
outside, then disengage. If he counter-disengages and step with a bind
into measure, re-disengage with haste, and thrust with a ricavazione
iutside fourth over his Blade. If he steps with counter-disengagement
not before him, then do him with a Ricavazione outside in fourth a
Feint with iron, and pay grand attention as to how he parries.
