\newpage

\scan{110}

\newpage

% Translation below

% i Secunda, så kommer du med din Klinga under hans / och utaff den
% linien din Arm / blifwer du betäckt. Fig. 21.

in Second, thus you come with your Blade under his, and from this line
in your Arm, you will be covered. Fig. 21.

\exercise{}
% Ett annat sätt. Wille din Advers', (när du ligger med jämbn Klinga)
% voltera Quart' uthan tempo, till din inwertes Kropp; Så stöt
% tillijka med honom en qvart a pie fermo öfwer hans swaga till hans
% Hals / eller förfall i det samma tempo med Secunda under hans
% Klinga. fig 21.

Another way. If your Adversary (when you lie with an even Blade) turn
Fourt oustide of tempo, to your inside Body. Then thrust with him a
fourth a pie fermo over his weak to his Throat, or fall in the same
tempo with Second under his Blade. fig. 21.

\exercise{}
% Stringera honom uthan till med porterer Klinga / wille han då gå
% igenom med ett insteg i den enge Mens', i mening at occupera din
% klinga innan till / eller stöta der en quarta, eller och wille
% förfalla under din klinga med en Secunda, så låt din motion aff ditt
% stringerande fara / och din udd uthan till siunka; Stöt så med
% Kropsens förskränckning a pie fermo under hans klinga / så at du
% kommer med din styrckia i hans swaga / så lärer du förhindra hans
% cavation. Du kan och denna Stötet giöra med en Passade.

Bind him outside with raised Blade, if he then goes through with an
instep in narrow Measure, intenting to occupy your Blade inside, or
thrust there a fourth, or wants to fal under your blade in Second,
then let the motion from your bind flow, and let your point sink
outside. Thrust then with e bend of the Body a pie fermo under his
blade, so that you come with your strong in his weak, thus you will
deny his disengage. You cabn also do this Thrust with a Pass.

\exercise{}
% Hafwer din Fiende stringerat din klinga utan till, så gack med din
% klinga igenom / och tag med din styrckia hans swaga / allenast med
% kropsens öfwerböjning; Med fötterne står du stilla; Caverar han i
% samma din rörelse / och wille giöra en contra-cavation, så giff
% grann acht uppå Tempo, i det ögneblek han wille stöta utan till
% öfwer din högra arm / at du träder med din wenstra foot på honom in
% / och stöter halfwa Tert, och halfwa Secunda contra-tempo till hans
% högre Bröst. Din wenstre hand förer du wäl fram / så kommer den
% högre deelen bättre sin koos utur presensa, hwar igenom du således
% med din wenstra Hand förbiuder och hindrar honom hans cavation.

If your Enemy has bound your blade outside, then go with your blade
through, and take with your strong his weak, with only the bending of
the body. Your feet remain still. If he disengages your motion, and
seeks to counter-disengage, then give close attention to Tempo, in the
instant he seeks to thrust outside over your right arm, that you step
in on him, and thrusts half Third, and half Second counter-tempo to
his right Chest. Your left hand bring well forward, so the right part
comes better out of presence, through which you thus with your left
Hand forbid and deny him his disengage.

\exercise{}
% Hafwer din Wederpart stringerat din klinga innan till / så föhr
% oförmärkt din högra foot i den enga Mens', gack i det samma med din
% Klinga under hans igenom / och med din styrckia öfwer hans swaga /
% så at du gifwer honom en öppning innan till åth ditt öfwerlijff / då
% låter han / i det samma du tager hans swaga / stöta innan till åth
% ditt öfwerlijff; Förfall då i samma tempo med Secunda under hans
% klinga. Fig. 21.

If your Opponent have bound your blade inside, then bring your right
foot unnoticed to narrow Measure, in the same moment go with your
Blade through under his, and with your string over his weak, so that
you give him an opening t your upper body, then he will, when you take
his weak, thrust inside to your upper body. Then fall in the same
tempo with Second under his blade. Fig. 21.

\exercise{}
% Hafwer din Advers' lägrat sig i secunda, så bint honom an med
% secund' utan till: Liggia i bägge allabala, så ryck något lijtet med
% din styrckia i hans swaga opföre / att han får tillfälle att stöta
% med secunda under din Klinga / giff gran acht i dett han Caverar
% under din Klinga / att du wänder din hand i Qvarta tillijka med
% honom / och stöter Qvarta med försänkter udd till hans underlijff.

If your Adversary settled in second, then bind him with second on the
outside. If you both stay [allabala???]\sidenote{Having searched and
asked, no one seems to know what this means, probably ``steady''?},
then twitch slightly with your strong in his weak upwards, so he has
an opportunity to thrust second under your Blade, pay good attention
as he Disengages under your blade, that you turn your hand in Fourth
with him, and thrust Fourth with lowered point to his groin.

\exercise{}
% Aft du af din Fiende innan till string, så stöt honom en Tert' utan till öfwer

If you are bound by your Enemy inside, then thrust him a Third outside over
