\newpage

\scan{077}

\newpage

% Translation below
\exercise{}
% Will din wederpart i samma Fint med Contrapostur (som skeer utan till
% med kroppen) förfalla och passera per secund' under Klingan till din
% inwertes Kropp att stöta / så låt i dett samma din udd uti Tert'
% siunka; Abaissera med kroppen / din Arm motte du intet sträckia
% aldeles uth / utan blifwa wijd hans swaga / giör fördenskull med
% kroppen en Scurtzo, så att din wenstre Axel kommer wähl fram / stöt
% i dett samma Tert' med ligation under hans Klinga med en
% passade. Fig. 8.
If your opponent in the Feint with Counterpose (done on the outside
wit the body) fall into and pass in second under the Blade, thrusting
to the inside of your body, then let immediately your point in Third'
sink. Lower your body, you must not extend your Arm fully, but stay by
his weak, for thsi do with the body a short\sidenote{Very tentative
translation of the original's ``Scurtzo''.}, so that your left
Shoulder is well extended, thrust immediately third with binding under
his Blade with a pass. Fig. 8.

%\execrcise{Huru förhängande Qvart' blifwer giord.}
\exercise{How overhanging Fourth is done.}
% Stringera honom innan till / och giör honom en fint utan till i
% Tert'; parerar han med jembd Klinga till sin högra sinkos, så
% falzera hans Klinga / och stöt innan till per Qvart' A. p. fermo
% Fig. 7.
Bind him on the inside, and do him a feint on the outside in Third, if
he parries with an even Blade to his right Side, then
false\sidenote{The original has ``falzera'', this could be a
``falcera''} his blade, and thrust on the inside in Fourth A. p. fermo
Fig. 7.

\exercise{}
% String' honom utan till / Caverar han och will åter igen innan till
% utan inträdning stringera, så Contra cavera och string' honom å uno
% utan till med ophäfning af din högre foot till den enga Mensuram; så
% fromt han åter går igenom/ så stöt Qvart' innan till med uddens
% försänkning / Klingan blifwer en wriden för än han har giordt sin
% Cavation; men du moste wähl blifwa med din styrka hos hans
% swaga. Fig. 7. 
Bind him on the outside, if he Diengages and wants to bind again on
the inside without entering, then Counterdisengage and bind him again
on the outside with lifting on your right foot to the short Measure;
the moment he again goes through, thrust Fourth on tye inside with a
lowering of the point, the Blade becomes twisted after he has done his
Disengage, but you need to stay well with your strong by his
weak. Fig. 7.

\exercise{}
% En Chiamata med förskrenckter Kropp. String' honom utan till / låt
% der på din udd wähl gå till din wänstra sinkos / så att du utan till
% blifwer aldeles blott / som i dett samma med kroppen
% fullkoml. skränkter / will han då i din swaga utan till instöta / så
% låt med en hast din wenstra Axel gå tillbaka igen / Cavera och
% voltera Qvart' till hans inwertes Kropp; Sätt din tåå i skränkning
% wähl uth / stöter han i din halfwa styrka så falzera hans Klinga och
% stöt Qvart' Fig 7.
An Invite wuth hunched Body. Bind him on the outside, let then your
point go well to your left side, so that you are completely open on
the outside, instantly bend your body, if he then seeks to thrust
oustide in your weak, then let with a haste your left Shoulder return,
disengage and rotate Fourth to the inside of his Body. Put your toe in
a bend well out, if he thrusts in your half strong then
false\sidenote{The original has ``falzera'', this could be a
mis-printed ``falcera'', it could be a wird related to the flattening
of leather or printed pages} his blade and thrust Fourth Fig. 7.
