\newpage

\scan{099}

\newpage

% när du förste gången geck öfwer hans Klinga / stöter du ända fram; Altså der han i din halfwa öfwerkastning/ skulle stöta / giör du en lijten parade. och der på en finte innan till åht hans Ansichte. Gif då gran acht huru han parerar.

when you initially go over his Blade, thrust all the way. Were he in your half overthrow, to thrust, you do a small parry. And then a small feint inside to his Face. At athta point, pay good attention to how he parries.

% \exercise{Emot öfwerkastning / och dem samme som stöta resolut / och icke med deporterer Klinga.}
\exercise{Against overthrows / and those who thrust robustly / and not with lowered\sidenote{This is a vert tentative translation of ``deporterer klinga''} Blade}
% Provocera din wederpart igenom Chiamater till stöten / och när han stöter / så bruka mäst dhe defensioner som sker uti twå tempi: En på slancerte och resolute stöter/ kan man säkert parera och sedan stöta/ emädan Fienden i dee samma sin effect intet kan mutera. Altså kan du när han innan till en qvarta aff frij stycken på dig wille stöta / under hans stöt Cavera. den samma utan till med qvarta parera. och träd i dett samma utan till hans högra sijda sinkos/ då du honom derstädes kan stöta/ eller med ett skärande öfwer hans wenstre kindbeen lädera.

Provoke your opponen by inviting to the tthust, and when he thrusts, use mostly those defences that happen in two tempos. Since on robust and strong thrusts, you can safely parry and then thrust, since the Enemy in these his effect cannot mutate. Thus you can when he on the inside to free parts wants thrust, you can Disengage under his thrust parry the same on the outside. And step immediately outside to his side, then you can him from there thrust or with a slice over his left cheekbone injure.

\exercise{}
% Altså och när han utan till stöter / Cavera på samma sätt under hans stöt / parera honom der innan till / och träd tilijka till hans wenstra sinkos / stöt derstädes med en öfwerkastning till hans högra Axel eller Ansichte eller och lädera honom med ett skärande öfwer hans högre kindben. Wilt du i hans stöt blifwa jämpt för honom / så parera den samme utan till med den wenstre tert' tillijka med en Scurtzio: Men utan

Similarly when he thrusts outside, Disengage in the same way under his thrust, parry him inside, and step to his left side, thrust then with an over-throw to his right Shoulder or Face or injure him with a slice over his right cheekbone. If you want in his thrust remain in line with him, parry it outside with left third, with a Scurtzio\sidenote{Translation yet unknown} But outside
