\newpage

\scan{045}

\newpage

% Translation below
% eller passera med secunda under hans Klinga bort: Stöter han med
% secunda till ditt Ansichte / så stöt per secundam tillijka åth hans
% öfwerdel eller Ansichte / Träder han innan till ett insteeg med sin
% swaga i din hela eller halfwa styrka / eller och stötte dijt in / så
% stöt du i dett samma innan till med ett insteeg af din högre eller
% wenstre fot per qvartam: Stötte han med en jembd eller förhängande
% Klinga till ditt under-lijff / så lått din udd i dett samma
% försiunka och voltera qvartam till hans högra höfft: Du kan och om
% du wilt / när han stöter med försänkter udd / träda in på honom /
% parera hans Klinga till hans högre sijda medföre / med din wenstra
% hand gripa i hans styrcka / och taga werjan utaf honom.
 
or pass with second under his Blade away. If  he thrusts with second
to your Face / thrust with second towards his upper body or Face. If
he steps in on the inside with his weak in your whole or half strong /
or thrusts there, thrust in that moment on the inside in fourth with a step in
with your right or left foot. If he thrusts with an even or dangling
Blade towards your groin, let your point in that moment sink, and
circle to fourth to his right hip. You can also if you want, when he
thrusts with a lowered point, step in on him, parry his Blade to his
right side, with your left hand grip his strong, and take the rapier
off him.

% Will han utan till string', Fintera eller stöta med styrkan i din
% swaga / så cavera i dett samma och stöt qvartam innan till hans
% kropp; Gieck han i en jämbd linea till ditt Ansichte / så förfall
% och stöt innan till per secundam.

If he binds on the outside, Feint, or thrust with the strong in your
weak, disengage immediately and thrust in fourth on the inside to his
body. If he goes in an even line to your Face, drop and thrust inside
in second.

% Stöter han i din halfwa swaga / så Cavera och stöt qvartam tillijka
% med falcering innan till hans kropp / wille han med ett insteeg
% stringera, Fintera eller stöta utan till med sin halfwa swaga /
% så wändt i secunda eller prima, då kommer du åter med din
% styrka till hans swaga / och stöt tillijka med honom contratempo,
% eller giör dett falske steget och stöt qvartam under hans Arm / utan
% rörelse hans Klinga . eller och angulera qvartam utan till öfwer
% hans Arm med dett falska steget.

If he thrusts in your half weak, Disengage and thrust in fourth also
with [falcera???]\sidenote{still not sure what this is, it may be
``bent'', cf. it \i{falcare}}, on the inside to his body, if he wants
with a step in bind, Feint or thrust on the outside with half his
weak. wait in second or first, then you will be again with your strong
in his weak, and thust him with counter-tempo, or do the false
step\sidenote{Very much not obvious what ``the false step'' is, at
this point}
and thrust in fourth under his Arm, without touching his Blade, or
angle in fourth outwards over his Arm with the false step.

% Skulle han i dett samma han stöter utan till i din halfwa Klinga /
% med qvarta giöra dett falske steget / så måtte du med en hast gå
% utur tertia och förwända i secunda / och i det samma gå innan till
% hans kropp; Du kan och wijd detta stöta tillijka contratempo med
% Secunda öfwer hans Arm; Men med kroppen måtte du wäl förfalla och
% handen wäl oplyffta.

Should he in that moment he thrusts on the outside in half your Blade,
in fourth do the false step, you must quickly go out of third and
turn in second, and simultaneously go inwards to his body; You can
also in this thrust counter-tempo with Second over his Arm. Wuth you
must well lower your body and well lift your hand.

% Gieck han utan till med ett insteeg af din swaga / till din halfwa
% eller hela styrka / eller stötte i dett samma / så gack tillijka med
% honom och stöt contratempo in öfwer hans Klinga; Stöter han utan
% till / under ditt kors / och du åst med dine fötter i linien hans
% Klinga / så stöt här jämpte under hans med ligation / eller
% försänck korset och stöt öfwer hans Klinga.

If he goes on the outside with a step in of your weak, to your half or
whole string, or thusts in that moment, so go with him and thrust
countertempo over his Blade. If he thrusts on the outside, under your
hilt, and your are with your feet in the line of his Blade, then
thrust evenly under his with a bind, or lower your hilt and thrust
over his Blade.

% Hwad beträffar de som som skiuta wärjan utur näfwan på deras
% Fiende / torde dett wäl lyckas emot de samma som låta sin
% Wederparts Klinga wara frij / och gifwa sig långt ifrån; Men emot de
% andre som weta wäl deras Fiende att occupera och string' / lära de
% intet stort uträtta / utan blifwa ofehlbart stötte och råkade.

As concerns those who shoot their rapier our of their fist towards
their Enemy, it should succeed well against those whoallow their
Opponent's Blade to be free, and give themselves far away. But against
those who know well to occupy and bind their Enemy, they will not
accomplish much, and will unerringly be thrusted and reached.
