\newpage

\scan{081}

\newpage

% Translation below
% till jern / der med han får åter tillfälle att string', när han då
% innan till med en påtryckning tillträder att string': Giör honom der
% på en meggia Cavation så att din Klinga kommer jempt under hans / då
% lärer han fela din Klinga / stöt i dett samma hastigt med meggia
% Cavation i hans intryckning per Tertiam till hans högre Axel Fig. 4.

iron, thus he again gets opportunity to bind, when he then on the
inside with a push steps to binds: Do him then a half Disengage, so
that your blade is even under his, then he will fail your Blade,
thrust instantly with half Disengage in his pressing in Third to his
right Shoulder Fig. 4.

% \exercise{Denna Fig. 9 med Wenstra Foten fram / utvisar en Passade}
\exercise{This Fig. 9 with the Left foot forward / demostrating a Passing}
% Ligger din Advers' långt i Ters' / så gack med din udd i prospective
% hans Kors / föhr din Klinga i Qvart' / träd i det samma du har fått
% Mens' med din högre fot i qvadro utur presensa hans udd / och behåll
% din wenstra sosp in Aria i lienien utaf din Fiendes Klinga / blif så
% med din Klinga i Qvarta jämpt under hans;
If your Adversary is long in Third, then go with your point towards is
Cross, bring your Blade into Fourth, step the instant you have Measure
with your right foot in fourth out of presence of his point, and keep
your left foot suspended in the air, in the line of your Enemy's
Blade, remain with your Blade in Fourth evenly under his.
%                                           Går han då prospective din
% Kropp effter / så sätt i det samma din Wänster Fot i sin förrige
% linie, något djupare in / och gack med din Klingan utan till hans;
% Enger han då med sin styrkia bak öfwer sig så cavera i det med
% Secunda under hans Klinga och passera den högre Foten in: Men der han
% (i det samma när du hafwer din Foot sosp. in Aria ) parerar din Klinga
% till sin högre sinkoos / så cedera och passera i det med wänstre
% Foten innan till på honom / till hans högre Bröst per
% Qvartam. Fig. 9.
If he then goes after your Body, then instantly set your Left Foot
back in its previous line, somewhat deeper in, and go with your Blade
outside his. If he narrows with his strong back over him then
disengage with Second under his Blade and pass the right Foot in. But
if he (in that you have your Foot suspended in the Air) parries your
Blade to his right lower, then retreat and pass in with your left Foot
on his inside. to his right Chest in Fourth. Fig. 9.

\exercise{}
% Ligger din Wederpart långt / och medh något ophäfwer udd / så gack
% honom an per Tertiam med udden i prospe hans kors intet mycket
% diupt: träd altså med wenstra foten i qvadri. och drag dig med
% Kroppen utur präsens hans udd / med Klingan blif i en jembd linea
% under hans / och håll den högre foten i samma linea sosp' in aria;
% Går han i dett med sin Klinga Prospect' din Kropp effetr / så sätt
% med en hast / den högre i
If your opponent is long, and with slightly elevated point, then go at
him in Third with the point seek his cross, but not too deep, step
thus with the left foot in fourth. And pull your body out of the
presence of his point, with your Blade be in an even line under his,
and hold the right foot in the same line suspended in the air. If he
in this goes seeking your Body with his Blade, then set with haste,
the right in
