\newpage

\scan{094}

\newpage

% Translation below

% \exercise{Flanconnade,}
% Har din Fiende stringerat dig innan till / så gack med din Klinga
% igenom och giff honom en falskt tempo eller chiamat till stöten åht
% din inwersz kropp / giff då grant acht i dett han stöter / att du
% giör en subtil parade, allenast med den fremsta leden så att din udd
% blifwer i praesensa hans kropp; Går han med sin hand i stöten
% temmeligen högt / och gifwer sig i sin högra sijda öppen / så
% förtryckt din hand i Qvarta och stöt den samme med en angulo under
% hand högre Arm / som denne Fig. 16. utwissar.
\exercise{Flanking}

If your Enemy bound you on the inside, then go through with your Blade
and give himm a false tempo or invite to thrust to your inside body,
pay good attention when he thrusts, that you make a subtle parry,
using only the frontmost joint so that your point is still in presence
to his body. If he goes high with his hand in the thrust, and opens
his right side, then press your hand into Fourth and thrust at an
angle under hist right Arm, as this Fig. 16. demonstrates.

\exercise{}
% Än mehr, Ligger din Fiende och har icke aldeles uhtsträkt sin Arm /
% utan lembnar sin högra sijda en blott / och du åst utaf honom
% string. så gack med en subtil cavation igenom med din styrka ia hans
% swaga / och stöt henne med qvart till hans högre sijda under högre
% Armen med en liten omwridning. Fig 16.

Furthermore, If your Enemy lies and has a not fully extended Arm, but
leaves his right side open, and you are bound by him, then go with a
subtle disengage through with your strong in his weak, and thrust the
blade in fourth to his right side under his Arm with a little
twist. Fig. 16.

\exercise{}
% Wijdare. hafwer din Fiende string dig utan till / heller du har
% string honom innan till / så stöt honom en finte innan till hans
% Ansichte; Parerar han med sin strykia öfwer sig till sin wenstra /
% och låtte udden siunka till din wenstra / så låter han ofwan till i
% sin högra sijda gifwa sig blott; Då förwändt med en hast din hand i
% full qvart' och stöt henne under hans Arm; Till större säkerheet så
% föhr din wenstra hand till förskänkning som Fig. 16, uhtwissar

Moving on. If your Enemy binds you on the outside, or you have bound him on the inside, thrust witha  feint inside to his Face. If he parries with his strong above to his left, he will give to his right side a high opening. Then turn your hand quickly into full fourth and thrust under his Arm. For greater safety move your left hand across as Fig. 16. demonstrates.

%\exercise{Huruledes Circulering och Ligering så wähl utan / som innan till / blifwa giorde.}
% Ligger din Advers' med en jembd och uhtsträckter Klinga för dig /
% och har string' dig innan till / så låht i dett samma din udd siunka
% / och din hand lyffter du opp med

\exercise{How circling and binding both outside and inside are done}

If your Adversary has an even and extended Blade in front of you, ad have bound you on the inside, in that instant let your point drop, and your hand you lift with
