\newpage

\scan{072}

\newpage

% Translation below

% till hans Ansichte / eller hwilket säkrare är åth hans högre Axel /
% med den högre foten sospeso in Aria: Kroppen och Armen blifwa wäl
% uthsträckte / parerar han öfwer sig till wänstra och gifwer tig
% öfwer Klingan ingen blott / så förfall i det samma och stöt per
% Tertiam i linien under hans Arm / wälbetächt a. pie fermo så at
% Handen blifwer i samma cito, och Udden allena giör effecten, parerar
% han intet effet finten, så stöt i det samma uti en motion per
% quartam, eller passera till hans högre Axel / caverar han under din
% fint, så wänt utur qvart i tertia, och stöt honom utan till hans
% högre Axel. Fig 4.
to his Face, or more safely to his right Shoulder, with the right foot
suspended in the Air. The Body and Arm well extended, if he parries
above to his left and give you no opening over the Blade, fall down
immediately and thrust in third under his Arm, well covered in a. pie
fermo so that the Hand remains in the same position, and the Point
alone makes the effect. if he does not parry after the feint, then
thrust immediately in one motion in fourth, or pass to his right
Shoulder, if he disengages under your feint, turn out of fourth into
third and thrust outside to his right shoulder. Fig. 4.

\exercise{}
% Giör honom finten som tillförende per quartam, tager han till sin
% wenstra / och giör sigh blott öfwer sin Klinga utan till / så cavera
% i samma cito, och stöt per Tertiam a. pie fermo och låt ditt
% Ansichte komma innan till. Fig 4.
Do him the feint as before in fourth, if he takes to his left, and
opens himself over his Blade on the outside, then disengage in the
same position, and thrust in third a. pie fermo and let your Face come
on the inside. Fig 4.

\exercise{The inside Feints with contra-position how they are doubled}
% Ligger din Wederpart långt / så bindt honom innan till och gå igenom
% / giör honom med Foten en battute, utan till en fint med lijtet
% Jern; Men intet diupt / parerar han / så cavera åter med en hast /
% och giör honom finten per quartam med contrapostur innan till / giff
% acht uppå huru han parerar.
If your Opponent lies long, then bind him on the inside and gp
through, do him wit the Foot a stomp, on the outside a feint with
small Iron. But not deep, if he parries, then disengage again with
haste, and do him the feint in fourth with counterpoise.

\exercise{}
% Stringera honom utan till / ligger han då stilla / så giör honom der
% med styrkan emot hans swaga en fint per Tertiuam eller Quartam till
% hans Ansichte eller högre Axel / parerar han öfwer sigh / så förfall
% och stöt med Secund under hans högra Arm till hans inwärtes Kropp /
% och så snart cavation är uthe / måtte stöten hafwa råkat. Du kant och
% stöta a. pie fermo eller passera, och giöra finten med blåtta
% Kropsens öfwerböjning / wil han intet parera effter finten, så
% förwändt henne gewsindt i en stöt per Tertiam til hans högre Axel /
% caverar han / så stöt Qvart med wenstre fotens passering till hans
% inwertz Kropp (Fig. 9.)
Bind him on the outside, if he remains still, then do him with the
strong against his weak a feint in Third or Fourth to his Face or
right Shoulder, if he parries above himself, then fall down and thrust
in Second under his right Arm to his inner Body, and as soon as the
disengage is done, the thrust must have reached. You can also thust
a. pie ferm or pass, and du the feint with a mere Bodily lean, if he
does not parry after the feint, turn your blade swiftly in a thrust in
Third to his right Shoulder, if he disengages, then thrust Fourth with
the left foot's passwing to his inwards body. (Fig. 9.)

\exercise{}
% Giör honom den finten utan till som tillförende / parerar han till
% sin högre i mening att bekomma din swaga / och will der med en hast
% gå för sig / tänkandes att stöta öfwer din swaga / så Cavera och
% voltera Qvart till hans inwertz kropp. Fig. 2.
Do him the feint on the outside as before, if he parries to his right
to posess your weak, and goes there with haste, intending to thrust
over your weak, then Disengage and circle in Fourth to his inwards
body. Fig 2.

\exercise{}
% Giör han paraden med wenstre Tertz; så stöt angulerat Qvartam utan
% till öfwer hans högre Arm in / att du tillika när du har råkat / kan
% wrijda hans Werja utur handen på honom. Fig. 15,
If he does the parry with Third left, thrust in Fourth on the outside over his right Arm, so that when you have hit can twist his Rapier out of his hand. Fig 15.

\exercise{}
% Lägg dig långt i Tertia med en jembd Klinga, wille han dig med ett
% insteg innan till String. så Cavera i dett samma förr än han rörer
% din Klinga / och stöt utan till per tertiam, så att när Cavation är
% ute / moste stöten hafwa råkar. (Fig. 4.)
Be in a long Third, even Blade. If he tries to Bind on the inside
with an instep, Disengage immediately before he touches your
Blade, thrust to the outside in third, so that when the Disengage
is done, the thrust must be done.

\exercise{}
% Sätter han i String. foten intet för sig / så giör honom i hans
% String. utan till findten per Tertiam med Contra postur och gif acht
% uppå huru han parerar, du kant och giöra Finten utan jern / eller
% och dubblera honom.
If he in Binding does not put his foot forward, do him in his Bind the
outside feint in Third with Counterpoise and pay attention to how he
parries, you can also do the Feint without iron, or double him.


\exercise{}
% Trycker din Adversarius utan till emot din swaga / och du har
% försummat ditt tempo med Cavation, så truck hårdt emot / och för än
% han får see sig före / så lått hans Klinga fara / och giör honom
% utan till finten i Tertia, stöt der på Secund, utan till under hans
% Klinga A. pie Fermo (Fig. 5.)
If your Opponent is pressing on your weak from the outside, and you have missed the tempo to Disengage, then press hard against, and before he has paid attention, then let his Blade go, and do him on the outside the feint in Third, ythrust then Second, on the outside under his blade a. pie fermo (Fig. 5.)

\exercise{How Second with some Lessons with a. pie fermo as with passages are done}
% Giör din Adversarius en finte per Qvartam, parerar han och drager till sig Klingan
Do your Adversary a feint in Fourth, if he parried and drag the Blade
