\newpage

\scan{074}

\newpage

% Translation below
%\exercise{Emot Mutering utur ett underläger}
% När din Fiende uthi Mutering med hela Armen och Klingan tillijka
% rörer sig; går du med din udd till hans högre Axel / framhävda i
% dett samma med honom och stöth Quart'. innan till. Item muterar han
% i ett underläger / så begif dig och i ett underläger med en jämbd
% Klinga / wille han åter gå i ett öfwerläger / så framträd du i dett
% samma / när han bryter Mensuren, och giör honom en Finte utan till
% hans Ansichte / cavera i dett samma / och stöt under hans Klinga. Du
% kant och när han muterar föllia hans Klinga / och gifwa grant acht
% uppå ifrå hwilken sijda han kommer / kommer han ifrå sin högra och
% går till sin wenstra / så stöter du Quart. men kommer han ifrå sin
% wenstra och går till sin högra / så stöter du Tertz. öfwer hans
% Klinga / men går han med sin Arm högt och rak så stöter du under
% Klingan. Fig 5.
\exercise{Against Mutation from a low position}
When your Enemy in Mutation with the whole Arm and Blade as well
moves, go with your point to his right Shoulder, throw foward at the
same time as him and thrust Fourth on the inside. If he mutates in a
low position, then go too to a low position with an even Blade, if he
wants to return to a high position, step forth immediately, whn he
breaks Measure, and do him a Feint on the outside to his Face,
disengage immediately, and thrust under his Blade. You can also when
he motates follow his Blade, and give grand attention on what side he
is coming, if he is coming from his right and goes to his left, then
thrust Fourth,., but if he comes from his left and goes to his right,
then thrust Third over his Blade, but if he goes with his Arm high and
straight then you trust under the Blade. Fig. 5.

\exercise{}
% Men der han (i dett samma att du stöter Tertz. öfwer halfwa styrkia)
% med sin Klinga far baköfwer / så passera med en hast fort och stöt
% Secund. uthan till under hans Klinga Fig. 5.

But when he (in the moment that you thrust in Third, over half strong)
with his Bloade fly backwards, then pass with a haste and thrust
Second, on the outside under his Blade. Fig. 5

\exercise{}
% Är du stringerat uthan till / så lått din udd med en hast
% nedersiuncka / gack i dett samma igen med din udd öfwer sin / och
% string'. honom hans Klinga innan till / så snart han går med sin
% Klinga igenom / så gif gran acht uppå Tempo i dett han går med sin
% Klinga igenom / då stöter du med Tertz. utan till öfwer hans halfwa
% styrkia af hans Klinga till hans högre Bröst / med ett insteg af
% högre foten. Fig. 4.

If you are bound on the outside, then let your point with haste drop,
go instantly again with your point over his, and bind his Blade on the
inside. As soon as he goes through with his blade, pay close attention
to tempo in that he goes with his Blade through, then you thrust with
Third on the outside over the half strong of hist Blade to his right
Chest, with an in-step of the right foot. Fig. 4.

\exercise{}
% Stringera din Adversarium utan til / caverar han / och wille innan
% till utan inträdning String- så cavera, och giör honom en fint utan
% till med Quart' till hans Ansichte, parerar han öfwer sig / så stöt
% innan till med hela Secund. a. p. fermo eller med Passaden, Fig. 5.

Bind your Adversary on the outside, if he disengages, and tries to
bind on the inside without instep, then disengage and do him a feint
on the outside with Fourth to his Face, if he parries upwards, then
thrust on the inside with whole Second a. p. fermo or with the
Passage. Fig. 5.

% \exercise{Huru man Stringering i ett medel- eller underläger bemöta skall.}
% Äst du innan till / och han wille stringera din Klinga utan till
% förutan inträdning / så giör honom i dett samma en Fint innan till /
% tager han med sin styrka effter din swaga / så cavera, i dett samma
% stöt Tertz. långt öfwer hans högre Arm / med uthstötning aff högre
% foten.
\exercise{How to meet Binding in a middle or lower position}
If you are on the inside, and he tries to bind your Blade on the
outside without stepping in, then do him in the same a Feint on the
inside, if he takes with his strong after your weak, then disengage,
and instantly thrust Thirrd long over his right Arm, with an outwards
thrust of the right foot.

\exercise{}
% Stringer din Fiende utan till / wille han i dett samma gå igenom
% och taga din Klinga innan till / så lått honom intet få röra wijd
% din Klinga / uthan i dett han är i motu, så giör honom en Fint i
% hans cavation utan till / i dett han då tager till sin högra sinkosz
% effter din Fint, så cavera och stöt Quart. innan till hans högre
% Bröst / och på samma sätt när du har stringerat din Fiende innan
% till Mutatis mutandis.

If your Enemy binds on the outside, if he attempts to go through and take your Blade on the inside, do not let him touch your Blade, but that when he is in motion, do him a Feint in his disengage on the outside, in that he then takes to the right after your Feint, disengage and thrust Fourth on the inside to his right Breast, and in the same way when you have bound your Enemy on the inside {\it Mutatis mutandis}\sidenote{roughly ``while doing the changes necessary to the situation''}

\exercise{}
% Har du lägrat dig långt / så fall i dett medel- eller under-läger;
% wille han då din Klinga innan till förföllia / så låts / i dett/ att
% du wilt innan till instöta; Men giff intet järn / parerar han / så
% stöt utan till / du kan och först ofwan och sedan nedan till
% fintera, så bringar du din Fiende i större Confusion, och står uti
% ditt behag hwart du wilt stöta öfwer eller under.

If yyou are in a long position, fall into a middle or low position, if
he then follows your blade to the inside, the let in that you want to
thrust on the inside. But do not give iron, if he parries, then thrust
on the outside. You can also above, then below, feint, thus you bring
your enemy into greter Confusion, and it is at your plesasure if you
want to trust above or below.
